---
title: "Natalie_GHTC_New"
author: "Natalie Smith-Gray"
date: "10/13/2019"
output: pdf_document
---

My Libraries
```{r}
library(MASS)
library(ISLR)
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(corrplot)
library(viridis)
library(WVPlots)
```

Upload the data
```{r}

consolidated_data <- read_csv("consolidated_data.csv")
head(consolidated_data)
attach(consolidated_data)
```

Natalie's Goals:  
I am going to look at:  
* Water into tank vs water out of the tank
* power vs water out of tank
* fux vs water out of tank
* temperature vs water out of tank
* wind speed vs temperature vs water out of tank

Therefore I need to get the following data from the dataframe:
* gal_in
* gal_out
* rate_in
* rate_out
* power
* pump_power?
* flux
* temp
**** wind_speed

```{r}
natalie_values <- consolidated_data %>%
  dplyr::select(gal_in, gal_out, in_rate, out_rate, power, pump_power, flux, temp, 'm/s Wind Speed') %>%
  na.omit() %>%
  cor()
round(natalie_values,3)
corrplot(natalie_values, order='hclust')
```

First lets look at water into tank vs water out of the tank
```{r}
time_series <- consolidated_data %>%
  dyplyr::select(time, out_rate, in_rate) %>%
  filter(time < )
plot(consolidated_data$time, 
     consolidated_data$in_rate, col = "green")
lines(consolidated_data$time,
      consolidated_data$out_rate, col = "red")
```
Now look at * fux vs water out of tank
```{r}
flux_out_rate <- consolidated_data %>%
  select(flux, out_rate) %>%
  na.omit()
plot(flux_out_rate)
lmr <- lm(out_rate ~ flux, flux*temp, data = consolidated_data)
summary(lmr)
plot(lmr)
```
Figure out the total daily use of water.
```{r}
water <- consolidated_data %>% 
  separate(time, c("date", "min"), sep = " ") %>%
  group_by(date) %>%
  mutate(sum_water_out = sum(out_rate, na.rm = TRUE)) %>%
  mutate(sum_water_in = sum(in_rate, na.rm = TRUE)) %>%
  filter(date < as.Date("2019-07-12")) %>%
  filter(date > as.Date("2019-06-27"))

#plot(water$date, water$sum_water_in)
ggplot(data = water, aes(x = date, y = sum_water_out)) +
  geom_point() +
  labs(x = "Date",
    y = "Water Used (gal)",
    title = "Amount of Water Used per Day") + theme(axis.text.x = element_text(angle = 45))
wawa <- water %>%
  select(temp, flux, sum_water_in, sum_water_out, date) %>%
  na.omit()

ggplot(wawa,)+ geom_point(aes(date, sum_water_in, col= "Water In")) + geom_point(aes(date, sum_water_out, col= "Water Out")) + theme(axis.text.x = element_text(angle = 45)) + labs(x = "Date", y = "Water Used (gal)", title = "Amount of Water Used per Day") + theme(plot.title = element_text(hjust = 0.5)) + theme(legend.title = element_blank())

hist(wawa$sum_water_in)
hist(wawa$sum_water_out)
pow_lux <- water %>%
  select(flux, power, temp, in_rate) %>%
  filter(flux < 700 & flux > 0 & power > 0 & in_rate > 0)

lem <- lm(power ~ flux, data = pow_lux)
summary(lem)
plot(pow_lux$flux, pow_lux$power)
abline(lem)

intercept = lem[["coefficients"]][["(Intercept)"]]
slope = lem[["coefficients"]][["flux"]]
sum_info <- summary(lem)

ggplot(pow_lux,) + geom_point(aes(flux, power, col=temp)) + scale_color_viridis() + labs(x = "Flux (W/m^2)", y = "Power (W)", title = "Solar Power Generation w/ Pump Running") + theme(plot.title = element_text(hjust = 0.5)) + geom_abline(slope = lem$coefficients[2], intercept = lem$coefficients[1]) + geom_text(aes(200, 750, label = paste("y =",slope, "x", " + ", intercept)))

plot(lem)

#https://winvector.github.io/WVPlots/reference/ScatterHist.html

set.seed(34903490)
y = water$in_rate
x = water$flux
frm = data.frame(x=x,y=y)
WVPlots::ScatterHist(frm, "x", "y",
  title= "Example Fit",
  smoothmethod = "gam",
  contour = TRUE)
######
testing <- water %>%
  select(power, current)
plot(water)
#plot(consolidated_data$time, 
     #consolidated_data$in_rate, col = "green")
#lines(consolidated_data$time,
      #consolidated_data$out_rate, col = "red")
```


```{r}
flux_temp <- consolidated_data %>%
  select(flux, temp, out_rate) %>%
  na.omit()
plot(flux_temp)
plot(temp,out_rate)
#lmr <- lm(out_rate ~ flux, data = consolidated_data)
#plot(lmr)
dfpv <- consolidated_data%>%
  select(out_rate, flux, temp)
ggplot(dfpv, aes(temp, out_rate, col=flux)) + geom_point() + scale_color_viridis()
#ggplot(wawa, aes(temp, sum_water_out, col=flux)) + geom_point() + scale_color_viridis()
```


```{r}
result <- consolidated_data %>%
  dplyr::select(current_raw, current_cal, voltage, flux, temp, in_rate, out_rate, power, in_rate, out_rate, pump_voltage, vdrop, pump_power, power_loss, 'W/m^2 Solar Radiation') %>%
  na.omit %>%
  cor()%>%
  corrplot(method="circle", order='hclust')
result
```

```{r}
dfpv <- consolidated_data%>%
  select(current_cal, voltage, flux, temp)

outliers <- which(dfpv$current_cal < -1)
dfpv$current_cal[outliers] <- 0
ggplot(dfpv, aes(voltage, current_cal, col=flux)) + geom_point() + scale_color_viridis()
```

# Wind Power
The objective of this section, is to explore if wind power is a feasible option for future projects in Panama. Some assumptions were made in the process and are labeled in the code below.
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
install.packages
# P = ?
rho = 1.225 # density of the air in kg/m^3 at sea level and 15 degrees C
pi = 3.1415926
d = 0.5 # wind mil diameter (m)
v <- consolidated_data$`m/s Wind Speed` # wind velocity
Wind_Pwr = (rho*pi*(d^2)*(v^3))/8
 
solar <- consolidated_data %>%
  select(time, power) %>%
  separate(time, c("date", "min"), sep = " ") %>%
  group_by(date) %>%
  mutate(solar_Whr = power*1/60) %>%
  mutate(energy_per_day = sum(solar_Whr, na.rm = TRUE)) %>%
  filter(date < as.Date("2019-07-12")) %>%
  filter(date > as.Date("2019-06-27"))
 
mean(solar$energy_per_day, na.rm = TRUE)
 
ggplot(solar,)+ geom_point(aes(date, energy_per_day, col = "Solar")) + theme(axis.text.x = element_text(angle = 45)) + labs(x = "Date", y = "Energy (Wh)", title = "Energy Per Day") + theme(plot.title = element_text(hjust = 0.5)) + theme(legend.title = element_blank())
```