---
title: "Solar Power Pump Project"
author: "Nathan Gray and Natalie Smith-Gray"
date: "12/12/2019"
output: pdf_document
---
## My Libraries
```{r}
#install.packages("tidyr")
library(MASS)
library(ISLR)
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(corrplot)
library(viridis)
library(reticulate)
use_python("C:\\ProgramData\\Anaconda3\\python.exe")
```
```{python}
from Photovoltaics import *
import numpy as np
```


## Upload the data
```{r}
consolidated_data <- read.csv("consolidated_data2.csv")
head(consolidated_data)
attach(consolidated_data)
```

# Correlation

## The first task is to create a scatter plot to look for any correlations between the data.
```{r}
scatter_plot <- consolidated_data %>%
  select(current_cal,voltage,flux,temp,in_rate,out_rate)
plot(scatter_plot)
```

## Correlation plots may be interesting to look at. 
```{r}
corr_plt_table <- consolidated_data %>%
  dplyr::select(current_cal,voltage,flux,temp,in_rate,out_rate,m.s.Wind.Speed) %>%
  na.omit() %>%
  cor()
round(corr_plt_table,3)
corrplot(corr_plt_table, order='hclust')
```

## Figure out the total daily use of water.
```{r}
water <- consolidated_data %>% 
  separate(time, c("date", "min"), sep = " ") %>%
  group_by(date) %>%
  mutate(sum_water_out = sum(out_rate, na.rm = TRUE)) %>%
  mutate(sum_water_in = sum(in_rate, na.rm = TRUE)) %>%
  filter(date < as.Date("2019-07-12")) %>%
  filter(date > as.Date("2019-06-27"))
```

## Look at the Water in and Water out
```{r}
#plot(water$date, water$sum_water_in)
ggplot(data = water, aes(x = date, y = sum_water_out)) +
  geom_point() +
  labs(x = "Date",
    y = "Water Used (gal)",
    title = "Amount of Water Used per Day") + theme(axis.text.x = element_text(angle = 45))
wawa <- water %>%
  select(temp, flux, sum_water_in, sum_water_out, date) %>%
  na.omit()

ggplot(wawa,) + 
  geom_segment(aes(x=date, xend=date, yend=sum_water_out, y=0), color='cyan4', size=4) + 
  geom_segment(aes(x=date, xend=date, yend=sum_water_in, y=sum_water_out), color='dark green', size=4) + 
  geom_point(aes(x=date, y=sum_water_in, color='Water Pumped'), size=3) + 
  geom_point(aes(x=date, y=sum_water_out, color="Water Used"), size=3) + 
  theme(axis.text.x = element_text(angle = 45)) + 
  labs(x = "Date", y = "Water Volume (gal)", title = "Water Pumped and Water Used Each Day") + 
  theme(plot.title = element_text(hjust = 0.5), legend.justification = 'right') + theme(legend.title = element_blank())
```

## Pump curve
```{r}
# Data fit for pump curve. Creates equation used for project.
df_flow <- consolidated_data%>%
  select(current_cal, voltage, vdrop, flux, temp, power, pump_power, in_rate)%>%
  filter(pump_power>30 & in_rate > 0 & in_rate < 14)
plot(df_flow$pump_power, df_flow$in_rate)

pumpfit = lm(in_rate ~ pump_power + I(pump_power^(1/2)), data=df_flow)
plot(pumpfit)
summary(pumpfit)
# -5.2733237 + -0.0066150(pump_power) + 0.6990381(pump_power)^(1/2)
# new calibration: -4.7752599 -0.0084187(pump_power) + 0.7501897(pump_power)^(1/2)
```

## Power vs Solar Irradiation
```{r}
pow_lux <- water %>%
  select(flux, power, temp, in_rate) %>%
  filter(flux < 718 & flux > 0 & power > 30 & in_rate > 0)

lem <- lm(power ~ flux, data = pow_lux)
summary(lem)

intercept = lem[["coefficients"]][["(Intercept)"]]
slope = lem[["coefficients"]][["flux"]]

ggplot(pow_lux,) + 
  geom_point(aes(flux, power), size=0.25, alpha=0.3) + 
  labs(x = "Flux (W/m^2)", y = "Power (W)", title = "Solar Power Generation w/ Pump Running") + theme(plot.title = element_text(hjust = 0.5)) + 
  geom_abline(slope = lem$coefficients[2], intercept = lem$coefficients[1]) + 
  geom_text(aes(200, 750, label = paste("y =",slope, "x", " + ", intercept))) +
  theme(axis.text.x = element_text(angle = 45))

plot(lem)
```

## Power Loss
```{r}
# Power Loss in wire versus power generated
dfpv <- consolidated_data%>%
  select(current_cal, voltage, flux, temp)

outliers <- which(dfpv$current_cal < -1)
dfpv$current_cal[outliers] <- 0
ggplot(dfpv, aes(voltage, current_cal, col=flux)) + geom_point() + scale_color_viridis()
```

## Do linear regression on flux vs water out of tank
```{r}
flux_out_rate <- consolidated_data %>%
  select(flux, out_rate) %>%
  na.omit()
plot(flux_out_rate)
lmr <- lm(out_rate ~ flux, flux*temp, data = consolidated_data)
summary(lmr)
plot(lmr)
```

## Look at the relationship between flux, temperature, and water usage
```{r}
flux_temp <- consolidated_data %>%
  select(flux, temp, out_rate) %>%
  na.omit()
plot(flux_temp)
plot(temp,out_rate)
dfpv <- consolidated_data%>%
  select(out_rate, flux, temp)
ggplot(dfpv, aes(temp, out_rate, col=flux)) + geom_point() + scale_color_viridis()
```

## Look at more cooralation plots. See if we missed any relationships
```{r}
result <- consolidated_data %>%
  dplyr::select(current_raw, current_cal, voltage, flux, temp, in_rate, out_rate, power, in_rate, out_rate, pump_voltage, vdrop, pump_power, power_loss, 'W.m.2.Solar.Radiation') %>%
  na.omit %>%
  cor()%>%
  corrplot(method="circle", order='hclust')
result
```
## Look at the relationship between current, flux, and voltage.
```{r}
dfpv <- consolidated_data%>%
  select(current_cal, voltage, flux, temp)
outliers <- which(dfpv$current_cal < -1)
dfpv$current_cal[outliers] <- 0
ggplot(dfpv, aes(voltage, current_cal, col=flux)) + geom_point() + scale_color_viridis()
```
```{python}

```

# Wind Power
The objective of this section, is to explore if wind power is a feasible option for future projects in Panama. Some assumptions were made in the process and are labeled in the code below.
```{r}
library(dplyr)
#install.packages
# P = ?
rho = 1.225 # density of the air in kg/m^3 at sea level and 15 degrees C
pi = 3.1415926
d = 15 # wind mil diameter (m)
v <- consolidated_data$m.s.Wind.Speed
#v <- consolidated_data$`m/s Wind Speed` # wind velocity
Wind_Pwr = (rho*pi*(d^2)*(v^3))/8

wind <- consolidated_data %>% 
  select(time,m.s.Wind.Speed, power) %>%
  separate(time, c("date", "min"), sep = " ") %>%
  group_by(date) %>%
  mutate(Wind_Pwr_W = (rho*pi*(d^2)*((m.s.Wind.Speed)^3))/8) %>%
  mutate(Wind_Pwr_Whr = Wind_Pwr_W*(15/60)) %>%
  mutate(sum_wind_pwr = sum(Wind_Pwr_Whr, na.rm = TRUE)) %>%
  mutate(solar_Whr = power/60) %>%
  mutate(power_per_day = sum(solar_Whr, na.rm = TRUE)) %>%
  filter(date < as.Date("2019-07-12")) %>%
  filter(date > as.Date("2019-06-27"))

ggplot(wind,)+ geom_point(aes(date, sum_wind_pwr, col = "Wind")) + geom_point(aes(date, power_per_day, col = "Solar")) + theme(axis.text.x = element_text(angle = 45)) + labs(x = "Date", y = "Power (W)", title = "Power Generation") + theme(plot.title = element_text(hjust = 0.5)) + theme(legend.title = element_blank())
```

#Linear Regression (NOT COMPLETED DUE TO LACK OF DATA...SEE NOTE AT THE END OF THIS SECTION OF CODE)
Regression model that models how much water out is used.
output: water out
input: power, temperature, solar irradiance, rainfall
model: linear regression, trees, pca to dimensional reduction

```{r}
# Organize the data for regression
ghtc_reg <- consolidated_data %>%
  select(time,power,W.m.2.Solar.Radiation,temp,out_rate,m.s.Gust.Speed,m.s.Wind.Speed,mm.Precipitation) %>%
  separate(time, c("date", "min"), sep = " ") %>%
  group_by(date) %>%
  fill(power,.direction = "down") %>%
  fill(W.m.2.Solar.Radiation,.direction = "down") %>%
  fill(temp,.direction = "down") %>%
  fill(W.m.2.Solar.Radiation, .direction = "up") %>%
  fill(m.s.Gust.Speed, .direction = "up") %>%
  fill(m.s.Wind.Speed, .direction = "up") %>%
  fill(mm.Precipitation, .direction = "up") %>%
  mutate(sum_water_out = sum(out_rate, na.rm = TRUE)) %>%
  mutate(avg_flux = sum(W.m.2.Solar.Radiation, na.rm = TRUE)/1440) %>%
  mutate(avg_gust = sum(m.s.Gust.Speed, na.rm = TRUE)/1440) %>%
  mutate(avg_wind = sum(m.s.Wind.Speed, na.rm = TRUE)/1440) %>%
  mutate(avg_precip = sum(mm.Precipitation, na.rm = TRUE)/1440) %>%
  mutate(avg_temp = sum(temp, na.rm = TRUE)/1440) %>%
  filter(date < as.Date("2019-07-12")) %>%
  filter(date > as.Date("2019-06-27"))

# Select only what you want for regression  
ghtc_reg <- ghtc_reg %>%
  select(sum_water_out,avg_flux,avg_gust,avg_precip,avg_wind,avg_temp)

# Correct for time difference
# Begin regression analysis
mlr <- lm(sum_water_out ~ avg_flux+avg_gust+avg_precip+avg_wind+avg_temp, data = ghtc_reg)
k <- step(mlr)
summary(k)
summary(mlr)
plot(mlr)

# Note that this linear regression was not completed due to the lack of data that was provided. However, it is all set up for when we do recieve more data. 
```

## Pump Curve
```{r}
# This chunck makes a model for the pump curve. We did not use this for the project since it produced bad results.
df_pump <- consolidated_data%>%
  select(current_cal, voltage, flux, temp, in_rate, pump_power, power, power_loss)%>%
  filter(pump_power > 20 & in_rate > 0 & in_rate < 14)

ggplot(df_pump, aes(pump_power, in_rate, col=temp)) + geom_point() + scale_color_viridis() + stat_smooth(method='lm', formula=y~poly(x, 3))
ggplot(df_pump,aes(x=power, y=in_rate )) + geom_point(size=0.25) + geom_density_2d() + labs(x="Power", y="Pump Rate", title="Pump Curve") +  theme(plot.title = element_text(hjust = 0.5)) + stat_smooth(method='lm', formula=y~poly(x, 3))
pump_curve_model = lm(df_pump$in_rate~poly(df_pump$pump_power, 3))
summary(pump_curve_model)
plot(pump_curve_model)
```

# Histograms
```{r}
hist_pwr <- consolidated_data %>%
  select(time,power) %>%
  separate(time, c("date", "min"), sep = " ") %>%
  group_by(min) %>%
  fill(power,.direction = "up") %>%
  mutate(avg_pwr = sum(power, na.rm = TRUE)/14) %>%
  filter(date < as.Date("2019-07-12")) %>%
  filter(date > as.Date("2019-06-27"))

average_day <- hist_pwr %>%
  filter(date == as.Date("2019-06-30"))
average_profile <- average_day$avg_pwr

E_avg = sum(average_profile)/60

normalized_profile = (average_profile)/E_avg
normalized_profile[normalized_profile < 0] <- 0
avg_energy_per_day = sum(normalized_profile)/60

hist(hist_pwr$avg_pwr)
plot(hist_pwr$avg_pwr)
plot(average_profile)
plot(normalized_profile)
```

# Optimization
```{r}
# We were having difficulties with the optimization program, so we trouble shooted by looking at it based on a day to day basis.

day_optimization <- consolidated_data %>%
  select(time,power,out_rate,gal_out,in_rate) %>%
  separate(time, c("date", "min"), sep = " ") %>%
  group_by(min) %>%
  filter(power >= 0)%>%
  fill(power,.direction = "down") %>%
  mutate(avg_pwr = sum(power, na.rm = TRUE)/14) %>%
  filter(date == as.Date("2019-07-3"))

water <- consolidated_data %>% 
  separate(time, c("date", "min"), sep = " ") %>%
  group_by(date) %>%
  mutate(sum_water_out = sum(out_rate, na.rm = TRUE)) %>%
  mutate(sum_water_in = sum(in_rate, na.rm = TRUE)) %>%
  filter(date < as.Date("2019-07-4")) %>%
  filter(date > as.Date("2019-07-2"))

Q_actual = 3477.6
# average water in per day = 3477.6 gal

E_avg = sum(average_profile)/60
normalized_profile = (average_profile)/E_avg
normalized_profile[normalized_profile < 0] <- 0
avg_energy_per_day = sum(normalized_profile)/60

E_1day = sum(day_optimization$power, na.rm = TRUE)/60
u_vector = (day_optimization$power)/E_1day
avg_energy_per_day = sum(u_vector)/60


# This is where the optimization starts. In the end we used the average energy profiles rather than just a day basis.
q_in <- function(E, is_size_1in) {
  u = normalized_profile                      # unit vector
  P = E*u
  sum = 0                                       # initializing a sum for q in for loop
  for(val in P) {
    if (is_size_1in){ #  1 inch
      q = (0.7105*(val^(1/2)))-(0.0021*val)-4.8752
    }else{ #  3/4 inch
      #q = (0.699*(val^(1/2)))-(0.006615*val)-5.273 # data fit
      q = 0.5918*val^(1/2) - 0.004*val - 3.6957      # theoretical (based on points eyeballed and located on pump curve) 9.34% error on Q, 12.33% error on E
      # new calibration: -4.7752599 -0.0084187(pump_power) + 0.7501897(pump_power)^(1/2)
      #q = -4.7752599 -0.0084187*(val) + 0.7501897*(val)^(1/2) # 17.43% error on E required, 12.68% error on Q
    }
    
    if(q < 0){
      q = 0
    }
    sum = sum + q
  }      
  print(sum)
  return_value = sum
}

g <- function(E, is_size_1in) {
  Q_min = Q_actual                                # min gal of water per day
  print(Q_min - q_in(E, is_size_1in))
  return_value = Q_min - q_in(E, is_size_1in)                    # g(E)
  #result_value = (s)
}

Q_validation = q_in(E_avg, FALSE)
print('% Q error')
print(abs(Q_validation-Q_actual)/Q_actual*100)

E_validation = 3736.54
#E_validation = 3519.1
q_in(E_validation, FALSE)
print('% E error')
print(abs(E_validation-E_avg)/E_avg*100)



E_1in = 2706.365                       # Energy required for a one-inch pipe
q_in(E_1in, TRUE)
E_3_4in = E_validation                  # Energy required for a three-quarters inch pipe

f <- function(E, size_1in){
  if (size_1in){
      p_ps = 3.57                         # Cost per 10 ft of one-inch pipe
    }else{
      p_ps = 2.44                         # Cost per 10 ft of one-inch pipe
    }
  p_pv = 1.25                            # 285 W panels at $1.25 per W
  n_pipes = 75                            # number of pipes
  sun_hours = 5                           # number of sun hours
  F_E = (p_pv*(E/sun_hours)) + (p_ps*n_pipes)  # f(E and pipe size)
}

print(f(E_1in,TRUE))
print(f(E_3_4in,FALSE))

```


Operation point comparison. Experimenting with different a_parameter values.
```{python} 

pvarray = Photovoltaics(voc=5 * 38.7, isc=9.42, vmp=5 * 32.1, imp=8.92, pmax=5 * 285, n=1.00, nseries=60 * 5)
pvarray.a_parameter = pvarray.a_parameter
temp = 273.15 + 27
flux = 672
v_meas = 137.77
i_meas = 5.94
varray = np.linspace(0, pvarray.voc, 100)
I = np.zeros(100)
for index, v in enumerate(varray):
	I[index] = pvarray.current(v, temp, flux)
plt.plot(varray, I, label='Model')
p = varray*I
pmax = max(p)
vmp_model = varray[np.argmax(p)]
imp_model = I[np.argmax(p)]
print(vmp_model)
plt.plot(vmp_model, imp_model, label='MPP', marker='o')
plt.plot(v_meas, i_meas, label='Operating Point', marker='x')
#plt.plot(139.374,7.560, label='actual', marker='x')
#plt.plot(139.374,7.560, label='actual', marker='x')
#plt.plot(139.374,7.2, label='actual', marker='x')
plt.axis([0,225, 0, 10])
plt.title('Array Model')
plt.legend()
plt.show()
print(pvarray.voltage(0, temp, flux))
```
```{r}
df_pvmodel <- consolidated_data%>%
  select(flux, temp, power, current_cal, voltage, pump_power, pump_voltage, in_rate)%>%
  filter(flux < 718 & flux > 0 & power > 30 & in_rate > 0)%>%
  mutate(model_power = py$pvarray$mpp_p(273.15 + temp, flux))
  
ggplot(df_pvmodel,) + 
  geom_point(aes(flux, power), size=0.25, alpha=0.3) + 
  geom_point(aes(flux, model_power), size=0.25, alpha=0.3, col='green') + 
  labs(x = "Flux (W/m^2)", y = "Power (W)", title = "Solar Power Generation w/ Pump Running") + theme(plot.title = element_text(hjust = 0.5)) + 
  geom_abline(slope = lem$coefficients[2], intercept = lem$coefficients[1]) + 
  geom_text(aes(200, 750, label = paste("y =",slope, "x", " + ", intercept))) +
  theme(axis.text.x = element_text(angle = 45))  
```


```{r}
df_pvmodel <- consolidated_data%>%
  select(flux, temp, power, current_cal, voltage, pump_power, pump_voltage, in_rate)%>%
  filter(flux < 718 & flux > 0 & power > 30 & in_rate > 0)
model_power = py$pvarray$mpp_p(c(305.15, 304.71), c(394.4, 384.4))
print(df_pvmodel$power[2000:2001])
273.15 + df_pvmodel$temp[2000:2001]
c(305.15, 304.71)
df_pvmodel$flux[2000:2001]
nparray_test <- py$np$array(c(394.4, 384.4))
py$pvarray$test()
```

